Canonical rings of rational divisors on projective space and Hirzebruch surfaces

### 2015-07-16 surfaces-write-up/surface-write-up.tex
edited projective non-1-point sections

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\todo{write more about hassett keel program}

@@

\begin{lem} \label{lem:proj-generators}
$R_D$ is generated in degrees up to $\sum_{i=1}^n \ell_i a_i.$
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \left \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n} \ell_i \bida_i = 0 \right \}.
\end{align}
We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e-i-proj}
	e_i = \left(\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots -\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j, -\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i \right)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Therefore, applying Proposition ~\ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

For the remainder of the proof, fix $i \in \{1,\ldots, n\}$. 
To complete the proof, it suffices to bound the degree of generation of the relations of the form $(z_i - \beta_i) \prod_{j=1}^n z_j^{c_j}$, by Equation ~\eqref{eqn:pm-relation-degree}.  We first associate $(z_i - \beta_i)\prod_{j=1}^n z_j^{c_j}$ with $(d, c_1, \ldots, c_n)$ where $d$ is its degree.  Let $\Sigma_i \subseteq \mathbb{Z}^{n+1}$ be the set of points of the form $(d, r_1, \ldots, r_n)$ satisfying $r_j \ge -d \alpha_j$ for all $j$ and $\sum_{j=1}^n r_j a_j = -a_i$.  
Then we see $\Sigma_i$ is contained in the $\mathbb{R}_{\ge 0}$-span of $e_1, \ldots, e_n$ (as defined in Equation \ref{defn:e-i-proj}) translated by by $\delta_i = (\frac{a_i}{\deg(D)}, -\frac{\alpha_1 a_1}{\deg(D)}, \ldots, - \frac{\alpha_n a_n}{\deg(D)})$.

Therefore, we can write any element of $\Sigma'$ as
\[
	\delta_i + \sum_{j=1}^n r_j e_j
\]
where $r_j \in \mathbb{R}$ for each $j$.

Whenever some $r_j \ge 1$, we can generate the relation at hand by one of lower degree by factoring out a copy of $\epsilon_j$ (see Equation ~\ref{eqn:epsilon-def-proj}).  Therefore, $\ker(\psi)$ is generated in degrees less than
\[
	\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

Therefore, since $\ker(\chi)$ and $\ker(\psi)$ are generated in degrees less than $2\sum_{i=1}^n \ell_i a_i$ and $\delta_i + \sum_{i=1}^n \ell_i a_i$ respectively, Lemma ~\ref{lem:composite-map} implies that $\ker(\phi)$ is generated in degrees less than
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]

\begin{thm}
\label{lem:proj-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes \di \bp^m$. Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $D_i = V(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]
with relations generated in degree at most
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{thm}

>>>>

\maketitle
\todo{make title shorter so it doesn't overlap with page numbers}
%%%%%%%%%%%%%%%%%%%%%%%%%%%% Introduction %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

@@

\begin{lem} \label{lem:proj-generators}
$R_D$ is generated in degrees up to $\sum_{i=1}^n \ell_i a_i.$
% and relations in degree at most  .\todo{insert numbers}
\end{lem}
\begin{proof}
Let 
\begin{align}\label{eqn:Sigma-def}
	\Sigma = \left \{(d, c_1, \ldots, c_n) \in \bz^{n+1} : c_i \geq - d \alpha_i, \; 1 \leq i \leq n, \text{ and } \sum_{i=1}^{n} \ell_i \bida_i = 0 \right \}.
\end{align}
\todo{Possibly take this definition out of the proof, and move it into the setup, since we want to reference it later.}
We note that $\Sigma$ has extremal rays given by the lattice points 
\begin{equation}\label{defn:e-i-proj}
	e_i = \left(\ell_i \bida_i, - \alpha_1 \ell_i \bida_i, \ldots -\alpha_{i-1} \ell_i \bida_i, \ell_i \sum_{j\ne i} \alpha_j \bida_j, -\alpha_{i+1} \ell_i \bida_i, \ldots, -\alpha_n, \ell_i \bida_i \right)
\end{equation}
for each $i\in \{1, \ldots n\}$.
Therefore, applying Proposition ~\ref{prop:cone-generation}, we see $R_D$ is generated in degrees less than
\[
	\sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

@@

For the remainder of the proof, fix $i \in \{1,\ldots, n\}$. 
To complete the proof, it suffices to bound the degree of generation of the relations of the form $(z_i - \beta_i) \prod_{j=1}^n z_j^{c_j}$, by 
Equation ~\eqref{eqn:pm-relation-degree}.  We first associate $(z_i - \beta_i)\prod_{j=1}^n z_j^{c_j}$ with $(d, c_1, \ldots, c_n) \in \Sigma,$ as defined in Equation ~\eqref{eqn:Sigma-def}.  
Let $\Sigma_i \subseteq \mathbb{Z}^{n+1}$ be the set of points of the form $(d, r_1, \ldots, r_n)$ satisfying $r_j \ge -d \alpha_j$ for all $j$ and $\sum_{j=1}^n r_j a_j = -a_i$.  Let 
\begin{align*}
	\delta_i = \left(\frac{a_i}{\deg(D)}, -\frac{\alpha_1 a_1}{\deg(D)}, \ldots, - \frac{\alpha_n a_n}{\deg(D)} \right).
\end{align*}
Then we see $\Sigma_i \subseteq \delta_i + \newspan_{\br_{\geq 0}}(e_1, \ldots, e_n)$ with $e_i$ as defined in Equation \ref{defn:e-i-proj}.
Therefore, we can write any element of $\Sigma_i$ as
\[
	\delta_i + \sum_{j=1}^n r_j e_j
\]
where $r_j \in \mathbb{R}$ for each $j$.

Whenever some $r_j \ge 1$, we can generate the relation $(z_i - \beta_i)\prod_{j=1}^n z_j^{c_j} = \epsilon_j \cdot h,$ where $\epsilon_j$ was defined in Equation ~\ref{eqn:epsilon-def-proj} and $h \in \Sigma_i$.  Therefore, $\ker(\psi)$ is generated in degrees less than
\[
	\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i.
\]
\end{proof}

Combining the above results, we obtain the following theorem:

\begin{thm}
\label{lem:proj-generators-relations}
Let $D = \sum_{i=1}^n \alpha_i D_i \in \bq \otimes \di \bp^m$. Let $f_i$ be homogeneous polynomials in $x_0, \ldots, x_m$ such that $D_i = V(f_i)$ for all $i$; suppose $\deg(f_i) = a_i$. 
Then $R_D$ is generated in degree at most 
\[
	\sum_{i=1}^n \ell_i a_i
\]
with relations generated in degree at most
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{thm}
\begin{proof}
The bound on degree of generation is precisely the content of Lemma ~\ref{lem:proj-generators}. It only remains to bound the degree of relations.
By ~\ref{lem:bound-ker-chi}, $\ker(\chi)$ is generated in degrees up to $2\sum_{i=1}^n \ell_i a_i$ and by Lemma ~\ref{lem:proj-relations-psi}, $\ker(\psi),$ is generated in degrees up to $\frac{\bida_i}{\deg(D)} + \sum_{i=1}^n \ell_i a_i.$ Consequently, Lemma ~\ref{lem:composite-map} implies that $\ker(\phi)$ is generated in degrees less than
\[
	\max \left(2 \sum_{i=1}^n \ell_i a_i, \frac{\max_{1\le i \le n}(\bida_i)}{\deg(D)} + \sum_{i=1}^n \ell_i a_i \right).
\]
\end{proof}
